---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# safejoin

<!-- badges: start -->

[![R build status](https://github.com/SamEdwardes/safejoin/workflows/R-CMD-check/badge.svg)](https://github.com/SamEdwardes/safejoin/actions) [![CRAN_Status_Badge](https://www.r-pkg.org/badges/version/safejoin)](https://cran.r-project.org/package=safejoin)

<!-- badges: end -->

## 🚧 Deprecation notice 🚧

As of `safejoin` version 0.2.0 the package has been deprecated.  As of version [`1.1.1`](https://dplyr.tidyverse.org/news/index.html#dplyr-111) dplyr has a `relationship` argument that provides the same functionality that `safejoin` was created for. See the dplyr docs [https://dplyr.tidyverse.org/reference/mutate-joins.html](https://dplyr.tidyverse.org/reference/mutate-joins.html) for complete details.

Please use `dplyr::left_join()` with the `relationship` argument instead.

## About

The goal of safejoin is to guarantee that when performing joins that extra rows are not added to your data. safejoin is a wrapper around the [`dplyr::left_join`](https://dplyr.tidyverse.org/reference/mutate-joins.html) function.

-   [Docs](https://safejoin-r.netlify.app/)
-   [GitHub](https://github.com/SamEdwardes/safejoin/)
-   [CRAN](https://CRAN.R-project.org/package=safejoin)

## Installation

You can install the released version of safejoin from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("safejoin")
```

Install development version from GitHub:

``` r
devtools::install_github("SamEdwardes/safejoin", ref = "dev")
```

## Example

Depending on your need safejoin can raise an error, a warning, or a message. By default safejoin will raise an error.

**Error**:

```{r example, error=TRUE}
library(safejoin)
x <- data.frame(key = c("a", "b"), value_x = c(1, 2))
y <- data.frame(key = c("a", "a"), value_y = c(1, 1))
safe_left_join(x, y, by = "key")
```

**Warning**:

```{r example_warning}
safe_left_join(x, y, by = "key", action="warning")
```

**Message**:

```{r example_message}
safe_left_join(x, y, by = "key", action="message")
```

When a join is "safe" `safe_left_join` will have the exact same behavior as [`dplyr::left_join`](https://dplyr.tidyverse.org/reference/mutate-joins.html).

```{r}
x <- data.frame(key = c("a", "b"), value_x = c(1, 2))
y <- data.frame(key = c("a", "b"), value_y = c(1, 1))
safe_left_join(x, y, by = "key")
```

## Other useful packages

There are other packages that help solve similar problems. Most notably <https://github.com/cynkra/dm> provides great features to treat data frames like a data base.

## Reference and Attribution

safejoin is created and maintained by Sam Edwardes.
